---
title: "Initial Korean Covid 19"
author: "Tyler Hamblin"
date: "10 October 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?



## Methods

According to the data found under kRdata we are trying to find out if there is a difference in the category Sex (male/female) in regards to COVID-19. Essentially we could prove if the virus is more harmful towards men or women based on the percentage of people from each group that pass away after contracting it. We can reach this conclusion by using Cat~Cat helper files from the CandP zip to help us decipher these statistics. 



## Graphical results - use ggplot commands to do this  


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In viewing the first chart we see that more women contracted COVID-19 compared to the amount of men included in the study. When looking at the second chart we see that a much higher percentage of men passed away from the virus. This second graph is important because with out it we would have assumed that the pandemic was worse for women. After comparing the percentage of deaths and survivals to the number of people in the count, there is a statistical suggestion that men fared worse. 


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This numerical chart confirms the results of the ggplot prior to it. In a total of 100 deaths, 36 of them would be women and the other 64 men. According to the second chunk of data it is over twice as likely for a man to die of COVID-19 than for a woman. 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Again the Fisher and Chisquare tests show that a man is at least twice (2.3x) as likely to die from the contraction of COVID-19 than a woman is. The p-values given from this code chunk are all consistent around the same range and although they are not as minuscule as values we have seen before, it still suggests statistical variation.  

## Conclusions

Our first step was loading and viewing the kRdata set and confirming that it was a categorical (result) described by a categorical (Sex). By using the Cat~Cat helper files from the CandP folder we found code chunks to run a ggplot that already suggested men were more prone to death from COVID-19. Next we ran a numerical test and inferential test in which the results and p-values confirmed the hypothesis that men who contract COVID-19 are more likely to die when compared to women, about twice as likely. We can confidently answer the original question that no sex is not independent of dying when looking at this particular data set. 


